2.3 Prob Model for PCR Positivity Estimation

### 2024-04-30 2_scripts/2.3 Prob Model for PCR pos.qmd
codes for sensiticity analysis

png(file="../4_output/figs/Fig.X1 AABB Conference.png", 
    width=2400, height=1800, res=300)
cowplot::plot_grid(p08.upper, p07 + theme(legend.position = c(0.95, 0.93),
                                          legend.background = element_blank()),
                   labels = c("","C"), ncol=1, rel_heights = c(1, 1))
dev.off()
```

@@

1. About the age dist in training data and CBS: The difference on age distribution might be a potential problem. Generally speaking, blood donors (CBS, FR-Leon data) is relatively older, while Covid-19 convalescent patients are younger.
```{r}
# The FR-Leon data
summary(dta_all[dta_all$study==1,]$age) # mean=46, median = 48
hist(dta_all[dta_all$study==1,]$age)

@@

>>>>

png(file="../4_output/figs/Fig.X1 AABB Conference.png", 
    width=2400, height=1800, res=300)
cowplot::plot_grid(p08.upper, p07 + theme(legend.position = c(0.95, 0.93),
                                          legend.background = element_blank()),
                   labels = c("","C"), ncol=1, rel_heights = c(1, 1))
dev.off()

# tiff format
tiff(file="../4_output/figs/Fig.X1 AABB Conference.tiff", 
    width=2400, height=1800, res=300, compression = "zip")
cowplot::plot_grid(p08.upper, p07 + theme(legend.position = c(0.95, 0.93),
                                          legend.background = element_blank()),
                   labels = c("","C"), ncol=1, rel_heights = c(1, 1))
dev.off()
```

@@

#### 4.1. About the age dist in training data and CBS: The difference on age distribution might be a potential problem. Generally speaking, blood donors (CBS, FR-Leon data) is relatively older, while Covid-19 convalescent patients are younger.
```{r}
# The FR-Leon data
summary(dta_all[dta_all$study==1,]$age) # mean=46, median = 48
hist(dta_all[dta_all$study==1,]$age)

@@

#### 4.2 Sensitivity Analysis

Both Alton and Yuan have address the importance of doing sensitivity analysis for the transformation models (Roche--Abbott). Originally we trained those two models with external matched data (Jamaica-Osaka-UK). We wonder if model would fit similarity with our own matched data in CBS.

#### 4.2.1 Using Jamaica-Osaka-UK data:
```{r}
# Roche-> Abbott
summary(lm01)

# Abbott-> Roche

summary(lm01r)
```
#### 4.2.2 Refit using CBS matched data
```{r}
CBSs2<-CBSs %>%
  .[!is.na(.$roche_n) & !is.na(.$abbott_n),] #CBS matched data

CBSs2$log_abbott_n<-log(CBSs2$abbott_n)
CBSs2$log_roche_n<-log(CBSs2$roche_n)

# clean inf values with log(0.01)
CBSs2$log_abbott_n[CBSs2$log_abbott_n==-Inf]<-log(0.01)

# lm05: Roche--> Abbott, using matched CBS
lm05<-glm(log_abbott_n ~ log_roche_n, 
          family = "gaussian",
          data = CBSs2)
summary(lm05)

# lm05r: Abbott--> Roche, using matched CBS
lm05r<-glm(log_roche_n~log_abbott_n , 
          family = "gaussian",
          data = CBSs2)
summary(lm05r)
# model 5r should be focused
```
#### Adjusting the weights of those with negative test results
```{r}
# add a indicator for positivity of two assays
CBSs2$ppnn[CBSs2$abbott_n > 1.4 & CBSs2$roche_n > 1]<-"pp"
CBSs2$ppnn[CBSs2$abbott_n > 1.4 & CBSs2$roche_n < 1]<-"np"
CBSs2$ppnn[CBSs2$abbott_n <=1.4 & CBSs2$roche_n < 1]<-"nn"
CBSs2$ppnn[CBSs2$abbott_n <=1.4 & CBSs2$roche_n > 1]<-"pn"

CBSs2$samplewt <- rep(1, length(CBSs2$pid))
CBSs2$samplewt[CBSs2$ppnn %in% c("np","nn")]<-0.01

# refit model 05r
lm05r.w<-glm(log_roche_n~log_abbott_n , 
          family = "gaussian", weights = samplewt,
          data = CBSs2)
summary(lm05r.w)
# after down-weighting, the betas became closer to 
```
